### 2014-11-10 tex/bisimulation.tex
Finished first draft of bisimulation

\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
\begin{align*}
\mathbb{HC}(p \rightharpoonup_{h} q) &= \\
 \bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\} 
\end{align*}
\end{definition}
The functional equivalence between heaps on a homomorphism and
constraint asserts a common structures in the two heaps under certain
conditions. The equivalence is used to related states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h}
q$ such that
\begin{equation}
\label{eqn:valid}
 \mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) ) 
\end{equation}
The represented relation is extended to sets of states $P$ and a single state $q$ as\footnote{The full proof argues $P \sqsubset Q$ but that $Q$ is
  always a singleton set containing $q$.}
$$
P \sqsubset q \Longleftrightarrow \forall p\ (p \sqsubset q \Rightarrow p \in P)
$$
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( (q \rsym q^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( (P \rsgse P^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
\end{definition}
As is typical, the functional reasons over a forward and backward
simulation, while the application to dissimilar states using a
meta-relation, $\rsgse$ is somewhat non-standard; though

>>>>

\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
\begin{align*}
\mathbb{HC}(p \rightharpoonup_{h} q) &= \\
 \bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\} 
\end{align*}
\end{definition}
The functional equivalence between heaps on a homomorphism and
constraint asserts a common structures in the two heaps under certain
conditions. The equivalence is used to related states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h}
q$ such that
\begin{equation}
\label{eqn:valid}
 \mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) ) 
\end{equation}
The represented relation is extended to sets of states $P$ and a single state $q$ as\footnote{The full proof argues $P \sqsubset Q$ but that $Q$ is
  always a singleton set containing $q$.}
$$
P \sqsubset q \Longleftrightarrow \forall p\ (p \sqsubset q \Rightarrow p \in P)
$$
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( (q \rsym q^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( (P \rsgse P^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
If $\sqsubset$ is a bisimulation, then the greatest fixed point of $F_\sim(\sqsubset)$ is the bisimilarity relation.
\end{definition}
As is typical, the functional reasons over a forward and backward
simulation, while the application to dissimilar states using a
meta-relation, $\rsgse$, is somewhat non-standard; though convenient
in this instance since it captures the notion of $q$ as a set of
states \cite{GSE:barbedbisimulation}.

\begin{lemma}[\textrm{F{\footnotesize IELD}} \textrm{A{\footnotesize CCESS}} preserves $\sqsubset\ \subseteq F_\sim(\sqsubset)$]
If $P \in 2^{S_\mathit{FA}}$ and $q \in S$ are such that $P \sqsubset q$, then $(P\ q)$ is in the functional associated to bisumlation.
\label{lem:access}
$$
\forall P \in 2^{S_\mathit{FA}}\ (P \sqsubset q \Rightarrow (P\ q) \in F_\sim(\sqsubset))
$$
\end{lemma}

\begin{proof}
Proof by contradiction: assume $P \sqsubset q \wedge (P\ q) \not\in F_\sim(\sqsubset)$.

\noindent\textbf{Sketch}: Choose any $P$ and $q$ at the field access continuation such
that $P \sqsubset q$. In the forward simulation
\eqref{eqn:BisimulationForwards}, for each $P^\prime$ such that $P
\rsgse P^\prime$, pick some $p^\prime \in P^\prime$. By definition $p \rgse
p^\prime$, $p \in P$, and $p \sqsubset q$. The proof shows the
existence of $q^\prime$ such that $q \rsym q^\prime$ and $p^\prime
\sqsubset q^\prime$ using the existing homomorphism in $p \sqsubset q$.

In the backward simulation \eqref{eqn:BisimulationBackwards}, for each
$q^\prime$ such that $q \rsym q^\prime$, the proof posits the
existences of some $p^\prime \sqsubset q^\prime$ and uses the
homomorphism in $p^\prime \sqsubset q^\prime$ to derive a $p$ such
that $p \rgse p^\prime$ and $p \in P$.

As the forward and backward simulations hold for any $p \in P$, $P
\sqsubset q$ must be in the functional, $F_\sim(\sqsubset)$, which is
a contradiction.
\end{proof}
Similar lemmas are proved for field write and equals reference.

\begin{theorem}
\label{th:bisim}
The relation $\sqsubset$ is a bisimulation.
\end{theorem}
\begin{proof}
By definition of the bisimilarity relation. All of the common rules
make no changes to the heap, so $P \subset q$ is included in the
functional. The other transition relations are supported by lemmas
such as \lemref{lem:access}.
\end{proof}

The notion $p \stackrel{n}{\rgse} p^\prime$ denotes that $p^\prime$ is
the $n$-step successor of $p$, if it exists, and $q
\stackrel{n}{\rsym} q^\prime$ is similarly defined. Completeness and
soundness are readily established by \thref{th:bisim} and inducting
over $n$.

\begin{corollary}[$\rsym$ is complete]
If $P \in 2^{S_\mathit{FA}}$ and $q \in S$ are such that $P \sqsubset q$ then
$$
\forall p^\prime\ (p \stackrel{n}{\rgse} p^\prime \Rightarrow \exists q^\prime\ (q \stackrel{n}{\rsym} q^\prime \wedge p^\prime \sqsubset q^\prime))
$$
\end{corollary}

\begin{corollary}[$\rsym$ is sound]
If $P \in 2^{S_\mathit{FA}}$ and $q \in S$ are such that $P \sqsubset q$ then
$$
\forall q^\prime\ (q \stackrel{n}{\rsym} q^\prime \Rightarrow \exists p^\prime\ (p \stackrel{n}{\rgse} p^\prime \wedge p^\prime \sqsubset q^\prime))
$$
\end{corollary}

The relation, $P \sqsubset q$, is readily established in the initial
state of a given program, as there is single initial state, $P_o =
\{p_o\}$, in any valid Javalite program. The initial state $\rsym$ is
then defined as $q_o = p_o$; $P_o \sqsubset q_o$ is trivially shown.
